---
title: "6_Analysis_degu-role"
author: "Annemarie van der Marel"
date: "2024-09-15"
output: html_document
---
What role do key individuals play?

In the presence of key individuals, degu’ social groups will show social stability in the form of increased connectivity (lower average path length and higher density), decreased aggression, increased foraging efficiency, high cohesion levels (larger diameter), and higher reproductive success and survival.

By group calculate 
- count
- diameter
- density
- average path length
- efficiency

- reproductive success  -> separate paper
- survival              -> separate paper


# load libraries
```{r setup}
#knitr::opts_chunk$set(echo = TRUE)
# R version 4.3.2 for all the packages to work

library(tidyverse)
library(job) # to run models in the background
library(gamlss)
library(ggpubr)
library(igraph)
library("assocInd")
library(betareg)
library(domstruc)
library(einet)
library(grid)
library(lme4)
library(brms)
library(ggraph)
library(muxViz)
```


# import data

dyad files

```{r fall X dyad info}

# proximity
#proxXdyad_fall <- read.csv( "results/proximityXdyad_fall24.csv")
prox_alldyads_fall <- read.csv( "../results/proximity_alldyads_fall24.csv") %>% 
  select(id1,id2, n, frequency)
n_distinct(prox_alldyads_fall$id1)
n_distinct(prox_alldyads_fall$id2)

# affiliative
#affXdyad_fall <- read.csv( "results/affiliativeXdyad_fall24.csv")
aff_alldyads_fall<-read.csv( "../results/affiliative_alldyads_fall24.csv")%>% 
  select(id1,id2, n, frequency)
n_distinct(aff_alldyads_fall$id1)
n_distinct(aff_alldyads_fall$id2)


# agonistic
#aggXdyad_fall <- read.csv( "results/agonisticXdyad_fall24.csv")
ag_alldyads_fall <- read.csv("../results/agonistic_alldyads_fall24.csv")%>% 
  select(id1,id2, n, frequency)
n_distinct(ag_alldyads_fall$id1)
n_distinct(ag_alldyads_fall$id2)

```




```{r}
# fall groups
rankings <- read.csv("../results/bordaranking_scaled_fall24_108degus.csv") %>% 
  select(-X, -X.1) %>% 
  arrange(BordaScore) %>% 
  mutate(rank= 1:n()) #bordaranking_fall90degus.csv")

selectdegusfall <- rankings$id


groups <- read.csv("../data/2024/groups_winter_24.csv") %>% 
  filter(id %in% selectdegusfall) %>% 
  dplyr::select(id, sex, burrow_July)
groups$id <- as.integer(groups$id)

groups$sex[groups$id=="7072"] <- "M"

n_distinct(groups$burrow_July)

```



# get data ready 
## helper functions

```{r}
#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}


overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model, type = "pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq / rdf
  pval <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
  c(chisq = Pearson.chisq, ratio = prat, rdf = rdf, p = pval)
}


normalize_sizes <- function(x, min_size = 5, max_size = 15) {
  (x - min(x)) / (max(x) - min(x)) * (max_size - min_size) + min_size
}

```



## group membership
```{r}
degulist24<- rankings %>% 
  left_join(groups) %>% 
  filter(!burrow_July %in% c(NA, "na")) %>% 
  rename(group = burrow_July)

n_distinct(degulist24$group)
```
Exclude groups of 1 degu
Add key degus

```{r}
groupingsselect <- degulist24 %>% 
  group_by(group) %>% 
  tally() %>% 
  filter(n!=1) 

groupsize <- degulist24 %>% 
  group_by(group) %>% 
  tally() %>% 
  rename(groupsize=n)

n_distinct(groupingsselect$group)
groupsselect <- as.integer(groupingsselect$group)

degulist24$group<-as.integer(degulist24$group)

degus24<- degulist24 %>% 
  filter(group %in% groupsselect) %>% 
  mutate(topkey= if_else(Borda<11, "key", "nonkey"),
         thkey=if_else(Borda<16, "key", "nonkey"),  # cutoff -> 15 key individuals
         qkey= if_else(BordaScore<quantile(BordaScore, probs = 0.25), "key", "nonkey"))

n_distinct(degus24$group)

# based on topten
groupXkey <- degus24 %>% 
  group_by(group, topkey) %>% 
  tally() 
table(groupXkey$topkey)

groups.with.key <- groupXkey$group[groupXkey$topkey=="key"]

# based on threshold
groupXthkey <- degus24 %>% 
  group_by(group, thkey) %>% 
  tally() 
table(groupXthkey$thkey)

groups.with.thkey <- groupXthkey$group[groupXthkey$thkey=="key"]

# based on lower quartile
groupXqkey <- degus24 %>% 
  group_by(group, qkey) %>% 
  tally() 

table(groupXqkey$qkey)
groups.with.qkey <- groupXqkey$group[groupXqkey$qkey=="key"]
```



```{r}

groups.key <- degus24 %>% 
    group_by(group) %>% slice(1) %>% 
  mutate(group_with_key=if_else(group %in% groups.with.key, "yes", "no"),
         group_with_thkey=if_else(group %in% groups.with.thkey, "yes", "no"),
         group_with_qkey=if_else(group %in% groups.with.qkey, "yes", "no")) %>% 
  select(group, group_with_key, group_with_thkey, group_with_qkey)
#groups.key$group <- as.character(groups.key$group)

sumgroups <- groups.key %>% 
  group_by(group_with_key) %>% 
  tally()

sumthgroups <- groups.key %>% 
  group_by(group_with_thkey) %>% 
  tally()


sumqgroups <- groups.key %>% 
  group_by(group_with_qkey) %>% 
  tally()

```

 
```{r}
degus24<- degus24 %>% 
  left_join(groups.key)
```

## dyadlist
of degus with ranking and group membership (total 92 ids)
```{r}
id <- degus24$id 

dyadlist <- expand.grid(id, id)
names(dyadlist) <- c("id1", "id2")
dyadlist <- subset(dyadlist, id1!=id2)

dyadlist1 <- expand.grid(id, id)
names(dyadlist1) <- c("actor", "subject")
dyadlist1 <- subset(dyadlist1, actor!=subject)

```

### top 10 
```{r}
# by id for associations
actor_group <- degus24 %>% 
  select(id, group, toptenkey, group_with_key) %>% 
  rename(id1= id, 
         actor_group= group, 
         actor_key=toptenkey, 
         actor_groupkey=group_with_key)

subject_group <- degus24 %>% 
  select(id, group, toptenkey, group_with_key) %>% 
  rename(id2= id, 
         subject_group= group, 
         subject_key=toptenkey,
         subject_groupkey=group_with_key)


```

```{r}
# by actor/subject for interactions

actor_group1 <- degus24 %>% 
  select(id, group, toptenkey, group_with_key) %>% 
  rename(actor= id, 
         actor_group= group, 
         actor_key=toptenkey, 
         actor_groupkey=group_with_key)

subject_group1 <- degus24 %>% 
  select(id, group, toptenkey, group_with_key) %>% 
  rename(subject= id, 
         subject_group= group, 
         subject_key=toptenkey,
         subject_groupkey=group_with_key)
```

### threshold
```{r}
# by id for associations
actor_group <- degus24 %>% 
  select(id, group, thkey, group_with_thkey) %>% 
  rename(id1= id, 
         actor_group= group, 
         actor_key=thkey, 
         actor_groupkey=group_with_thkey)

subject_group <- degus24 %>% 
  select(id, group, thkey, group_with_thkey) %>% 
  rename(id2= id, 
         subject_group= group, 
         subject_key=thkey,
         subject_groupkey=group_with_thkey)


```

```{r}
# by actor/subject for interactions

actor_group1 <- degus24 %>% 
  select(id, group, thkey, group_with_thkey) %>% 
  rename(actor= id, 
         actor_group= group, 
         actor_key=thkey, 
         actor_groupkey=group_with_thkey)

subject_group1 <- degus24 %>% 
  select(id, group, thkey, group_with_thkey) %>% 
  rename(subject= id, 
         subject_group= group, 
         subject_key=thkey,
         subject_groupkey=group_with_thkey)
```




#calculate group-level network metrics per group instead of the complete network


### proximity
```{r proximity}
start.time <- Sys.time()
run.s <- as.character(unique(groups.key$group))
replicates <- unique(groupingsselect$group)

#make empty dataframe to fill
networkmetricsXgroup_prox <- data.frame(group=character(), 
                                    count=numeric(),
                                       
                                       ec.value=numeric(),
                                       diameter=numeric(), 
                                       density=numeric(),
                                       avg_path_length=numeric(), 
                                       #connectedness, 
                                       efficiency=numeric(),
                                       ei=numeric(), 
                                       prunk.pu=numeric(), 
                                       prunk.dyads=numeric(),
                                       network=character()
                                   )  

plot_prox_list <- list()  # Initialize empty list for network graphs


for(run in 1:length(unique(groups.key$group))){
  run.code <- run.s[run]
  group <-run.code
  print(run.code)

dyadXgroup <- degus24 %>% 
  filter(group==run.s[run]) %>% # check burrow 16, 70, 72,75, 99 #
  select(id) 
dyadXgroup_ids <- dyadXgroup$id

dyadXgroup_list <- expand.grid(dyadXgroup_ids, dyadXgroup_ids)
names(dyadXgroup_list) <- c("id1", "id2")
dyadXgroup_list <- subset(dyadXgroup_list, id1!=id2)


 # proximity
prox_alldyads_fall$id1 <- as.integer(prox_alldyads_fall$id1 )

prox_groupdyads <- right_join(prox_alldyads_fall, dyadXgroup_list ) 
 #fill newly-merged data with 0's where no interactions
 prox_groupdyads[is.na(prox_groupdyads)] <- 0

   #print a check
  check <- length(prox_groupdyads$id1)
  
         # Check if count (number of rows) in subset_data is 0
  if (sum(prox_groupdyads$n) == 0) {
    # Skip this iteration if no data
    next  # 'next' jumps to the next iteration of the loop
  }

# matrix
  mx.prox.group <- reshape2::dcast(prox_groupdyads, id1~id2, value.var="frequency") #n for counts vs frequency controlled for observation bias
  mx.prox.group[is.na(mx.prox.group)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx.prox.group <- matrix.please(mx.prox.group)

# network
  graph.prox.group <- graph_from_adjacency_matrix(mx.prox.group, "undirected", 
                                                  weighted=T, diag=F)
  g <- graph.prox.group
  

# add vertex/node attribute
nodes <- degus24 %>% 
  filter(group==run.s[run])
nodes$size <- normalize_sizes(nodes$rank, min_size = 16, max_size = 9)

V(g)$name <- nodes$id
V(g)$rank <- nodes$size 
V(g)$sex <- nodes$sex
#V(g)$key <- nodes$toptenkey
V(g)$key <- nodes$thkey

  
# visualise network
  p <- ggraph(g, layout = "fr") +
    
    # edges
    geom_edge_link(aes(edge_width = weight), alpha = 0.8) +
    scale_edge_width(range = c(0.2, 1)) +  
    
    # nodes
    geom_node_point(aes(shape= key,
                       fill = sex,
                       color = "white", 
                       size=rank)) + 
    scale_shape_manual(name = "key", 
                       values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                       labels = c( "key", "nonkey" )) +
    scale_fill_manual(name = "sex",
                      values=  c("#f03b20", "#2c7fb8"),
                      labels = c( "F", "M" )) + 
    scale_color_manual(values="white") +
    #scale_size_continuous(range = c(min_size, max_size)) +
    
    ggtitle(paste("Group", run))+
    theme_void() +
    theme(legend.position = "none",
          title = element_text(size=8))
 
   plot_prox_list[[run]] <- p  # Store plot in list
  
  

# network metrics
  count= sum(prox_groupdyads$n)  
  
  eigenvector <- (igraph::eigen_centrality(g, directed = TRUE,
                                      weights = E(g)$weight))
  #ec <- eigenvector$vector
  ec.value <- eigenvector$value
  
  diameter = diameter(g)
  
  density = edge_density(g)
  
  avg_path_length = average.path.length(g) # mean_distance = mean_distance((g)), -> calculates the same
  
  #connectedness = vertex.connectivity(g)
  
  ei = einet::effective_information(g, effectiveness = FALSE)
  efficiency <- ei/log2(length(dyadXgroup_ids))

  
  #Find proportion unknown relationships, a measure of sparseness
  prunk <- EloRating::prunk(mx.prox.group)
  prunk.pu <- as.numeric(prunk[1])
  prunk.dyads <- as.numeric(prunk[2]) 
  
metrics_prox_groupXgroup <- cbind.data.frame(group, 
                                             count,
                                       ec.value,
                                       diameter, 
                                       density,
                                       avg_path_length, 
                                       #connectedness, 
                                       efficiency,ei, 
                                       prunk.pu, prunk.dyads
                                       ) %>% 
  mutate(network="proximity")

  # POOL ALL DATA

  networkmetricsXgroup_prox <- rbind.data.frame(networkmetricsXgroup_prox, 
                                           metrics_prox_groupXgroup)
}

end.time <- Sys.time()

#Time to run:
end.time - start.time






```

### affiliative
```{r proximity}
start.time <- Sys.time()
run.s <- as.character(unique(groups.key$group))
replicates <- unique(groupingsselect$group)

#make empty dataframe to fill
networkmetricsXgroup_aff <- data.frame(group=character(), 
                                    count=numeric(),
                                       
                                       ec.value=numeric(),
                                       diameter=numeric(), 
                                       density=numeric(),
                                       avg_path_length=numeric(), 
                                       #connectedness, 
                                       efficiency=numeric(),
                                       ei=numeric(), 
                                       prunk.pu=numeric(), 
                                       prunk.dyads=numeric(),
                                       network=character()
                                   )  #,
                                       #linearity=numeric(),
                                       #steepness=numeric(),
                                       #transi.ttri=numeric()) 

# Initialize empty list for network graphs
plot_aff_list <- list() 


for(run in 1:length(unique(groups.key$group))){
  run.code <- run.s[run]
  group <-run.code
  print(run.code)

dyadXgroup <- degus24 %>% 
  filter(group==run.s[run]) %>% 
  select(id) 
dyadXgroup_ids <- dyadXgroup$id

dyadXgroup_list <- expand.grid(dyadXgroup_ids, dyadXgroup_ids)
names(dyadXgroup_list) <- c("id1", "id2")
dyadXgroup_list <- subset(dyadXgroup_list, id1!=id2)


# affiliative
aff_groupdyads <- right_join(aff_alldyads_fall, dyadXgroup_list ) 
 #fill newly-merged data with 0's where no interactions
 aff_groupdyads[is.na(aff_groupdyads)] <- 0
 
 
        # Check if count (number of rows) in subset_data is 0
  if (sum(aff_groupdyads$n) == 0) {
    # Skip this iteration if no data
    next  # 'next' jumps to the next iteration of the loop
  }

# matrix  (rows are actors, columns are subjects)
  mx.aff.group <- reshape2::dcast(aff_groupdyads, id1~id2, value.var="frequency") #head(ref.behavior1.mx)
  mx.aff.group [is.na(mx.aff.group )] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx.aff.group  <- matrix.please(mx.aff.group )
  
  
# network
  graph.aff.group  <- graph_from_adjacency_matrix(mx.aff.group , "directed", weighted=T, diag=F) 
  #plot(graph.aff.group )

# network metrics
   g <- graph.aff.group 
   
   # add vertex/node attribute
nodes <- degus24 %>% 
  filter(group==run.s[run])
nodes$size <- normalize_sizes(nodes$rank, min_size = 16, max_size = 9)

V(g)$name <- nodes$id
V(g)$rank <- nodes$size 
V(g)$sex <- nodes$sex
#V(g)$key <- nodes$toptenkey
V(g)$key <- nodes$thkey

  
# visualise network
  a <- ggraph(g, layout = "fr") +
      
    # edges
    geom_edge_fan(aes(edge_width = weight, 
                    alpha = 0.6 # ..index..
                    ), 
                strength = 0.05) +
    scale_edge_width(range = c(0.2, 1)) +  
    
    # nodes
    geom_node_point(aes(shape= key,
                       fill = sex,
                       color = "white", 
                       size=rank)) + 
    scale_shape_manual(name = "key", 
                       values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                       labels = c( "key", "nonkey" )) +
    scale_fill_manual(name = "sex",
                      values=  c("#f03b20", "#2c7fb8"),
                      labels = c( "F", "M" )) + 
    scale_color_manual(values="white") +
    #scale_size_continuous(range = c(min_size, max_size)) +
    
    ggtitle(paste("Group", run))+
    theme_void() +
    theme(legend.position = "none",
          title = element_text(size=10))
 
   plot_aff_list[[run]] <- a  # Store plot in list


# individual metrics
  degree <- igraph::degree(g, mode = c("out"))
  
  strength <- strength(g, mode = ("out"), weights = E(g)$weight)
  
  closeness <- igraph::closeness(g)
  
  between <- igraph::betweenness(g, 
                                directed = TRUE,
                                weights = E(g)$weight)
  
  eigenvector <- (igraph::eigen_centrality(g, directed = TRUE,
                                      weights = E(g)$weight))
  ec <- eigenvector$vector
  ec.value <- eigenvector$value
  
  # page rank
  pr.list <- igraph::page_rank(g, 
                          directed = TRUE, 
                          weights = E(g)$weight)
  prv <- pr.list$vector
  pr <- prv/max(prv)
  
  
  metrics_aff_indXgroup <-  cbind.data.frame(group,
                                             
                           #int.ids, #z2.degus, fall_degus_rc
                           degree,
                           strength, 
                           between, 
                           closeness,
                           ec,
                           pr) %>% 
    mutate(network="affiliative")
  
  
# network metrics
   count= sum(aff_groupdyads$n) 
  
  diameter = diameter(g)
  
  density = edge_density(g)
  
  avg_path_length = average.path.length(g) # mean_distance = mean_distance((g)), -> calculates the same
  
  #connectedness = vertex.connectivity(g)
  
  ei = einet::effective_information(g, effectiveness = FALSE)
  efficiency <- ei/log2(length(selectdegusfall))

  
  #Find proportion unknown relationships, a measure of sparseness
  prunk <- EloRating::prunk(mx.aff.group )
  prunk.pu <- as.numeric(prunk[1])
  prunk.dyads <- as.numeric(prunk[2]) 
  
metrics_aff_groupXgroup <- cbind.data.frame(group, 
                                            count,
                                       ec.value,
                                       diameter, 
                                       density,
                                       avg_path_length, 
                                       #connectedness, 
                                       efficiency,ei, 
                                       prunk.pu, prunk.dyads
                                       ) %>% 
  mutate(network="affiliative")



  # POOL ALL DATA
  networkmetricsXgroup_aff <- rbind.data.frame(networkmetricsXgroup_aff, 
                                           metrics_aff_groupXgroup)
}

end.time <- Sys.time()

#Time to run:
end.time - start.time






```


### agonistic
```{r agonistic}
start.time <- Sys.time()
run.s <- as.character(unique(groups.key$group))
replicates <- unique(groupingsselect$group)

#make empty dataframe to fill
networkmetricsXgroup_agg <- data.frame(group=character(), 
                                    count=numeric(),
                                       
                                       ec.value=numeric(),
                                       diameter=numeric(), 
                                       density=numeric(),
                                       avg_path_length=numeric(), 
                                       #connectedness, 
                                       efficiency=numeric(),
                                       ei=numeric(), 
                                       prunk.pu=numeric(), 
                                       prunk.dyads=numeric(),
                                       network=character()
                                   )  #,
                                       #linearity=numeric(),
                                       #steepness=numeric(),
                                       #transi.ttri=numeric()) 


plot_agg_list <- list() 


for(run in 1:length(unique(groups.key$group))){
  run.code <- run.s[run]
  group <-run.code
  print(run.code)

dyadXgroup <- degus24 %>% 
  filter(group==run.s[run]) %>% # check burrow 16, 70, 72,75, 99 #
  select(id) 
dyadXgroup_ids <- dyadXgroup$id

dyadXgroup_list <- expand.grid(dyadXgroup_ids, dyadXgroup_ids)
names(dyadXgroup_list) <- c("id1", "id2")
dyadXgroup_list <- subset(dyadXgroup_list, id1!=id2)


# aggression


agg_groupdyads <- right_join(ag_alldyads_fall, dyadXgroup_list ) 
 #fill newly-merged data with 0's where no interactions
 agg_groupdyads[is.na(agg_groupdyads)] <- 0

# matrix  (rows are actors, columns are subjects)
  mx.agg.group <- reshape2::dcast(agg_groupdyads, id1~id2, value.var="frequency") #head(ref.behavior1.mx)
  mx.agg.group [is.na(mx.agg.group )] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx.agg.group  <- matrix.please(mx.agg.group )
  
  
# network
  graph.agg.group  <- graph_from_adjacency_matrix(mx.agg.group , "directed", weighted=T, diag=F) 
  #plot(graph.agg.group )

# network metrics
   g <- graph.agg.group 
   
      count= sum(agg_groupdyads$n) 
      

       # Check if count (number of rows) in subset_data is 0
  if (sum(agg_groupdyads$n) == 0) {
    # Skip this iteration if no data
    next  # 'next' jumps to the next iteration of the loop
  }

# Code here runs only if subset_data has rows (count > 0)

 # set node attributes   
nodes <- degus24 %>% 
  filter(group==run.s[run])
nodes$size <- normalize_sizes(nodes$rank, min_size = 16, max_size = 9)

V(g)$name <- nodes$id
V(g)$rank <- nodes$size 
V(g)$sex <- nodes$sex
#V(g)$key <- nodes$toptenkey
V(g)$key <- nodes$thkey

  
# visualise network
  f <- ggraph(g, layout = "fr") +
      
    # edges
    geom_edge_fan(aes(edge_width = weight, 
                    alpha = 0.6 # ..index..
                    ), 
                strength = 0.05) +
    scale_edge_width(range = c(0.2, 1)) +  
    
    # nodes
    geom_node_point(aes(shape= key,
                       fill = sex,
                       color = "white", 
                       size=rank)) + 
    scale_shape_manual(name = "key", 
                       values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                       labels = c( "key", "nonkey" )) +
    scale_fill_manual(name = "sex",
                      values=  c("#f03b20", "#2c7fb8"),
                      labels = c( "F", "M" )) + 
    scale_color_manual(values="white") +
    #scale_size_continuous(range = c(min_size, max_size)) +
    
    ggtitle(paste("Group", run))+
    theme_void() +
    theme(legend.position = "none",
          title = element_text(size=10))
 
   plot_agg_list[[run]] <- f  # Store plot in list
   
# individual metrics

  dom_ec <- dom_ec(mx.agg.group) # dominance rank  
  
  degree <- igraph::degree(g, mode = c("out"))
  
  strength <- igraph::strength(g, mode = ("out"), weights = E(g)$weight)
  
  closeness <- igraph::closeness(g)
  
  between <- igraph::betweenness(g, 
                                directed = TRUE,
                                weights = E(g)$weight)
  
  eigenvector <- (igraph::eigen_centrality(g, directed = TRUE,
                                      weights = E(g)$weight))
  ec <- eigenvector$vector
  ec.value <- eigenvector$value
  
  # page rank
  pr.list <- igraph::page_rank(g, 
                          directed = TRUE, 
                          weights = E(g)$weight)
  prv <- pr.list$vector
  pr <- prv/max(prv)
  
  
  metrics_agg_indXgroup <-  cbind.data.frame(group, 
                           #int.ids, 
                           dom_ec,
                           degree,
                           strength, 
                           between, 
                           closeness,
                           ec,
                           pr) %>% 
    mutate(network="agonistic")
  
  
# network metrics

   
  diameter = diameter(g)
  
  density = edge_density(g)
  
  avg_path_length = average.path.length(g) # mean_distance = mean_distance((g)), -> calculates the same
  
  #connectedness = vertex.connectivity(g)
  
  ei = einet::effective_information(g, effectiveness = FALSE)
  efficiency <- ei/log2(length(selectdegusfall))

  
  #Find proportion unknown relationships, a measure of sparseness
  prunk <- EloRating::prunk(mx.agg.group)
  prunk.pu <- as.numeric(prunk[1])
  prunk.dyads <- as.numeric(prunk[2]) 
  
  # # dominance metrics
  # #Find linearity
  # linearity <- EloRating::h.index(mx.agg.group, loops = 1000)
  # linearity <- linearity[3,2]
  # 
  # #Find steepness
  # steepness <- steepness::getStp(mx.agg.group, method="Dij")
  # 
  # #Triangle transitivity
  # transi <- EloRating::transitivity(mx.agg.group, runs = 1000)
  # transi.Pt <- transi[1]  # proportion of transitive triads
  # transi.ttri <- transi[2]  # triangle transitivity

# job::job({  
#   # dominance pattern
#   focus <- dom_focus(mx.agg.group)
#   position <- dom_position(mx.agg.group)
#   fp <- cbind.data.frame(focus, position)
#   colnames(fp) <- c("focus", "position")
#   
#   #Compute blur models
#   blur <- dom_make_blur_data(mx.agg.group)
#   
#   #Find strategy
#   dompattern <- dom_categorize_strategy(data=fp, blur_data=blur)  
# })
#   
metrics_agg_groupXgroup <- cbind.data.frame(group, 
                                            count,
                                       ec.value,
                                       diameter, 
                                       density,
                                       avg_path_length, 
                                       #connectedness, 
                                       efficiency,ei, 
                                       prunk.pu, prunk.dyads) %>% 
                                       #linearity,
                                       #steepness,
                                       #transi.ttri) %>% 
                                       #dompattern) %>% 
  mutate(network="agonistic")
  
 

  
  # POOL ALL DATA
  # run.sums <- rbind.data.frame(metrics_prox_groupXgroup, 
  #                              metrics_aff_groupXgroup, 
  #                              metrics_agg_groupXgroup )
  networkmetricsXgroup_agg <- rbind.data.frame(networkmetricsXgroup_agg, 
                                           metrics_agg_groupXgroup )
}

end.time <- Sys.time()

#Time to run:
end.time - start.time


```

### multilayer

```{r}

start.time <- Sys.time()
run.s <- as.character(unique(groups.key$group))
replicates <- unique(groupingsselect$group)

#make empty dataframe to fill
multiplexmetricsXgroup <- data.frame(group=character(), 
                                    m.apl=numeric(),
                                    cluster=numeric(),
                                    overlap=numeric(),
                                    diameter=numeric(),
                                    density=numeric(),
                                    density_pool=numeric()
                                   )  

plot_multi_list <- list()  # Initialize empty list for network graphs


for(run in 1:length(unique(groups.key$group))){
  run.code <- run.s[run]
  group <-run.code
  print(run.code)

dyadXgroup <- degus24 %>% 
  filter(group==run.s[run]) %>% # check burrow 16, 70, 72,75, 99 #
  select(id) 
dyadXgroup_ids <- dyadXgroup$id

dyadXgroup_list <- expand.grid(dyadXgroup_ids, dyadXgroup_ids)
names(dyadXgroup_list) <- c("id1", "id2")
dyadXgroup_list <- subset(dyadXgroup_list, id1!=id2)


 #### proximity -------------

prox_groupdyads <- right_join(prox_alldyads_fall, dyadXgroup_list ) 
 #fill newly-merged data with 0's where no interactions
 prox_groupdyads[is.na(prox_groupdyads)] <- 0

   #print a check
  check <- length(prox_groupdyads$id1)
  
         # Check if count (number of rows) in subset_data is 0
  if (sum(prox_groupdyads$n) == 0) {
    # Skip this iteration if no data
    next  # 'next' jumps to the next iteration of the loop
  }

# matrix
  mx.prox.group <- reshape2::dcast(prox_groupdyads, id1~id2, value.var="frequency") #n for counts vs frequency controlled for observation bias
  mx.prox.group[is.na(mx.prox.group)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx.prox.group <- matrix.please(mx.prox.group)

  # network
  graph.prox.group <- graph_from_adjacency_matrix(mx.prox.group, "undirected", 
                                                  weighted=T, diag=F)

  #### affiliative --------------------
aff_groupdyads <- right_join(aff_alldyads_fall, dyadXgroup_list ) 
 #fill newly-merged data with 0's where no interactions
 aff_groupdyads[is.na(aff_groupdyads)] <- 0
 
 
        # Check if count (number of rows) in subset_data is 0
  if (sum(aff_groupdyads$n) == 0) {
    # Skip this iteration if no data
    next  # 'next' jumps to the next iteration of the loop
  }

# matrix  (rows are actors, columns are subjects)
  mx.aff.group <- reshape2::dcast(aff_groupdyads, id1~id2, value.var="frequency") #head(ref.behavior1.mx)
  mx.aff.group [is.na(mx.aff.group )] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx.aff.group  <- matrix.please(mx.aff.group )
  
  
# network
  graph.aff.group  <- graph_from_adjacency_matrix(mx.aff.group , "directed", weighted=T, diag=F) 
  #plot(graph.aff.group )
  
#### aggression ------------------------
agg_groupdyads <- right_join(ag_alldyads_fall, dyadXgroup_list ) 
 #fill newly-merged data with 0's where no interactions
 agg_groupdyads[is.na(agg_groupdyads)] <- 0

# matrix  (rows are actors, columns are subjects)
  mx.agg.group <- reshape2::dcast(agg_groupdyads, id1~id2, value.var="frequency") #head(ref.behavior1.mx)
  mx.agg.group [is.na(mx.agg.group )] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx.agg.group  <- matrix.please(mx.agg.group )
  
  
# network
  graph.agg.group  <- graph_from_adjacency_matrix(mx.agg.group , "directed", weighted=T, diag=F) 
  #plot(graph.agg.group )


  
#### multilayer network ---------------
df3list <- list( prox_groupdyads, aff_groupdyads, agg_groupdyads) 
g3list <- list(graph.prox.group, graph.aff.group, graph.agg.group)  
node3tensor <- list( mx.prox.group, mx.aff.group, mx.agg.group) # NodesTensor, list of matrices
layer3tensor <- diagR(c(1, 1), 3, 1) + diagR(c(1, 1), 3, -1)# LayerTensor
l = 3 # Layers, number of layers 
n = length(dyadXgroup$id) # Nodes, number of nodes 

# build graph
m <- BuildSupraAdjacencyMatrixFromEdgeColoredMatrices(NodesTensor = node3tensor, 
                                                      LayerTensor = layer3tensor,
                                                      Layers = l,
                                                      Nodes = n)

# aggregate
m.pool<- muxViz::GetAggregateMatrix(node3tensor, Layers = l, Nodes = n)
graph.pool <- GetAggregateNetworkFromNetworkList(g3list)

# plot
layercolors <- c("#332288", "#44AA99", "#CC6677") #"#88CCEE", 

p_m <- plot_multiplex(g3list,
  layercolors,
  edge.colors = "grey",
  node.colors = "auto",
  node.size.values = 0.5,
  node.alpha = 1,
  edge.alpha = 1,
  layout = "fr",
  show.legend = FALSE)

plot_multi_list[[run]] <- p_m  # Store plot in list


# multilayer metrics (SupraAdjacencyMatrix, Layers, Nodes)

path <-	GetMultiPathStatistics(m, l, n)
m.apl <- path$avg.path.length

cluster<- GetAverageGlobalClustering(m, l, n) # The result ranges from 0 (no clustering) to 1 (maximal clustering/cohesion). 

overlap <- GetAverageGlobalOverlapping(m, l, n) #High overlap = cohesive multiplex structure; low overlap = more segregated/interlayer variation.

assort <- GetInterAssortativityTensor(m, l, n, isDirected= T, Type = "TT")
assort$InterSpearman

# diameter
g_multi <- graph_from_adjacency_matrix(m, mode = "undirected", weighted = NULL)
diameter <- diameter(g_multi, directed = FALSE, unconnected = TRUE)

# density 
num_nodes_multi <- nrow(m)
num_edges_multi <- sum(m[upper.tri(m)] != 0)
max_edges_multi <- num_nodes_multi * (num_nodes_multi - 1) / 2

density <- num_edges_multi / max_edges_multi

# aggregate metrics
# Density for aggregate network (monoplex)
num_nodes <- nrow(m.pool)

# Number of edges for undirected network without loops
num_edges <- sum(m.pool[upper.tri(m.pool)] != 0)

# Maximum possible edges 
max_edges <- num_nodes * (num_nodes - 1) / 2

# Calculate density
density_pool <- num_edges / max_edges
print(density_pool)
  
  
# path.pool <-	GetMultiPathStatistics(m.pool, l, n)
# m.pool.apl <- path.pool$avg.path.length
# 
# cluster.pool <- GetAverageGlobalClustering(m.pool, l, n) # SupraAdjacencyMatrix, Layers, Nodes
# 
# overlap.pool <- GetAverageGlobalOverlapping(m.pool, l, n)


#### combine -------------

metricsXgroup <-  cbind.data.frame(group, 
                                   diameter, 
                                   density,
                                    m.apl,
                                    cluster,
                                    overlap,
                                    density_pool) %>% 
  mutate(network="multilayer")
  

 multiplexmetricsXgroup <- rbind.data.frame(multiplexmetricsXgroup, 
                                            metricsXgroup )
}

end.time <- Sys.time()

#Time to run:
end.time - start.time


```

### combine




to create the graphs I had to remove the groups without any interactions: 
```{r}
 networkmetricsXgroup24 <-  bind_rows(networkmetricsXgroup_prox,
                                      networkmetricsXgroup_aff,
                                      networkmetricsXgroup_agg,
                                      multiplexmetricsXgroup) %>% 
  mutate(year="2024") 
write.csv(networkmetricsXgroup24,
          "../results/networkmetricsXgroup_fall24_excludesnulls_v1.csv")


```

combine with sex

```{r}
keysex24 <- degulist24 %>% 
  filter(rank<16) %>% 
  select(sex, group) %>% 
  rename(sex.keyID=sex)

keysex24$group <- as.factor(keysex24$group)
keysex24$sex <- as.factor(keysex24$sex)
networkmetricsXgroup24$group <- as.factor(networkmetricsXgroups24$group)


metricsXkey.sex <- networkmetricsXgroup24 %>% 
  left_join(keysex24) %>% 
  drop_na(sex) 

metricsXkey.sex24 <- metricsXkey.sex

write.csv(metricsXkey.sex,
          "../results/networkmetricsXgroupXsex_fall24.csv")
```



This is run on all groups, no matter whether they had no interactions. 
```{r}

n_distinct(networkmetricsXgroup$group)

head(networkmetricsXgroup)
networkmetricsXgroup24 <- networkmetricsXgroup 


sum<-  left_join(groups.key, networkmetricsXgroup) %>% 
  filter(network=="proximity") %>% 
  group_by(group_with_key) %>% 
  tally()
  
n_distinct(networkmetricsXgroup24$group)

write.csv(networkmetricsXgroup24,
          "results/networkmetricsXgroup_fall24_23groups.csv")
```

# comparing group-level metrics
## compare the network metrics of groups with and without key individuals 

```{r}

selectnetwork <- "proximity" # "proximity" "affiliative" "agonistic"

networkmetricsXgroup24$group <- as.integer(networkmetricsXgroup24$group) 
networkmetricsXgroupXkey <- networkmetricsXgroup24 %>% 
  left_join(groups.key)


# run.code <- unique(networkmetricsXgroupXkey$network)
# replicates <- 3
# 
# start.time <- Sys.time()
# 
# observed.summary<- data.frame(network=character(), 
#                               diameter.obs=numeric(),
#                               density.obs=numeric(), 
#                               avg_path_length.obs=numeric(), 
#                               efficiency.obs=numeric()) 
# 
# 
# for(run in 1:replicates){
#   network <- run.code[run]
#   print(network)
# 


```


### observed proximity
```{r proximity}

networkmetricsXgroup_prox$group <- as.integer(networkmetricsXgroup_prox$group)

proxXkey <- networkmetricsXgroup_prox %>% 
  left_join(groups.key)   
  

ref.data <- proxXkey 

```

```{r number of associations}

range(proxXkey$count)

# normally distributed?
ggplot(proxXkey, aes(x=count)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed



# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(proxXkey$count)


m <- lm(count ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(proxXkey$count)
fitdistrplus::descdist(proxXkey$count, discrete = T) # normal distribution


ggplot(proxXkey, aes(x=group_with_key, y=count)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$diameter, proxXkey$withkey)
wilcox.test(count~group_with_key, data=ref.data)

# Calculate effects
#count.obs <- coefficients(glm(count ~ group_with_key, data = ref.data, family = poisson))[2]
count.obs <- coefficients(glm(count ~ group_with_thkey, data = ref.data, family = poisson))[2]
#count.obs <- coefficients(glm(count ~ group_with_qkey, data = ref.data, family = poisson))[2]

#summary(glm(count ~ group_with_key, data = ref.data, family = poisson))

```


```{r connectivity}

range(proxXkey$diameter)

# normally distributed?
ggplot(proxXkey, aes(x=diameter)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

ref.data <- proxXkey %>% 
  filter(diameter<1.5)

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(proxXkey$diameter)


m <- lm(diameter ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(proxXkey$diameter)
fitdistrplus::descdist(proxXkey$diameter, discrete = F) # normal distribution


ggplot(proxXkey, aes(x=group_with_key, y=diameter)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$diameter, proxXkey$withkey)
wilcox.test(diameter~group_with_key, data=ref.data)

# Calculate effects
#diameter.obs <- coefficients(lm(diameter ~ group_with_key, data = ref.data))[2]
diameter.obs <- coefficients(lm(diameter ~ group_with_thkey, data = ref.data))[2]
#diameter.obs <- coefficients(lm(diameter ~ group_with_qkey, data = ref.data))[2]


```




``` {r density}

####### density  #####

range(proxXkey$density)

# normally distributed?
ggplot(proxXkey, aes(x=density)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(proxXkey$density)


m <- lm(density ~ group_with_key, data = proxXkey)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(proxXkey$density)
fitdistrplus::descdist(proxXkey$density, discrete = F) # beta distribution

ggplot(proxXkey, aes(x=group_with_key, y=density)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$density, proxXkey$group_with_key)
wilcox.test(density~group_with_key, data=proxXkey)

# Calculate effects
proxXkey$density[proxXkey$density==1.0]<-0.9999

#density.obs <- coefficients(betareg(density ~ group_with_key, data = proxXkey))[2]
density.obs <- coefficients(betareg(density ~ group_with_thkey, data = proxXkey))[2]
#density.obs <- coefficients(betareg(density ~ group_with_qkey, data = proxXkey))[2]
```

```{r information transfer}

###### average path length ######
ref.data <- proxXkey %>% 
  filter(avg_path_length<1)

range(ref.data$avg_path_length, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=avg_path_length)) +
  geom_histogram()  # very right skewed

  # shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$avg_path_length)


m <- lm(avg_path_length ~ group_with_key, data = ref.data)
#m<- glm(avg_path_length ~ withkey, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$avg_path_length)
#fitdistrplus::descdist(ref.data$avg_path_length, discrete = F) # beta distribution

ggplot(ref.data, aes(x=group_with_key, y=avg_path_length)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$avg_path_length, proxXkey$group_with_key)
wilcox.test(avg_path_length~group_with_key, data=ref.data)

m1<- glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse"))
m2 <- lm(avg_path_length ~ group_with_key, data = ref.data)

performance::compare_performance(m1,m2, rank = TRUE, verbose = FALSE)
performance::check_model(m1)
performance::check_model(m2)


#avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse")))[2]
avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_thkey, data = ref.data, family=Gamma("inverse")))[2]
#avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_qkey, data = ref.data, family=Gamma("inverse")))[2]

#avg_path_length.obs <- coefficients(lm(avg_path_length ~ group_with_key, data = ref.data))[2]


```

```{r efficiency}


##### efficiency #####
range(ref.data$efficiency, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=efficiency)) +
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$efficiency)


m <- lm(efficiency ~ group_with_key, data = ref.data)
#m<- glm(efficiency~ group_with_key, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
  fitdistrplus::plotdist(ref.data$efficiency)
fitdistrplus::descdist(ref.data$efficiency, discrete = F)


# Calculate effects
ref.data$efficiency[ref.data$efficiency==0.0]<-0.0001
ref.data$efficiency[ref.data$efficiency==1.0]<-0.9999

#efficiency.obs <- coefficients(betareg(efficiency ~ group_with_key, data = ref.data))[2]
efficiency.obs <- coefficients(betareg(efficiency ~ group_with_thkey, data = ref.data))[2]
#efficiency.obs <- coefficients(betareg(efficiency ~ group_with_qkey, data = ref.data))[2]

ggplot(ref.data, aes(x=group_with_key, y=efficiency)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$efficiency~proxXkey$withkey)
wilcox.test(efficiency~group_with_key, data=ref.data)
```



```{r}
# combine coeffs
network <- "proximity"
prox.obs<- cbind.data.frame(network, count.obs, diameter.obs,density.obs, avg_path_length.obs, efficiency.obs)
```


in proximity networks, we found no difference in network metrics between groups with and groups without key individuals 






### observed affiliative
```{r observed affiliative }


ref.data <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") 


```

```{r number of associations}

range(ref.data$count)

# normally distributed?
ggplot(ref.data, aes(x=count)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed


# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$count)


m <- lm(count ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$count)
fitdistrplus::descdist(ref.data$count, discrete = T) # normal distribution


ggplot(ref.data, aes(x=group_with_key, y=count)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$diameter, proxXkey$withkey)
wilcox.test(count~group_with_key, data=ref.data)
# Calculate effects
count.obs <- coefficients(glm(count ~ group_with_thkey, data = ref.data, family = poisson))[2]
#summary(glm(count ~ group_with_key, data = ref.data, family = poisson))

```

```{r connectivity}
###### connectivity

range(ref.data$diameter)

# normally distributed?
ggplot(ref.data, aes(x=diameter)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$diameter)


m <- lm(diameter ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$diameter)
fitdistrplus::descdist(ref.data$diameter, discrete = F) # beta distribution


ggplot(ref.data, aes(x=group_with_key, y=diameter)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$diameter, proxXkey$withkey)
wilcox.test(diameter~group_with_key, data=ref.data)
# Calculate effects
diameter.obs <- coefficients(lm(diameter ~ group_with_thkey, data = ref.data))[2]

```

```{r}
####### density  #####

range(ref.data$density)

# normally distributed?
ggplot(ref.data, aes(x=density)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$density)


m <- lm(density ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$density)
fitdistrplus::descdist(ref.data$density, discrete = F) # beta distribution

ggplot(ref.data, aes(x=group_with_key, y=density)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$density, proxXkey$withkey)
wilcox.test(density~group_with_key, data=ref.data)

# Calculate effects
ref.data$density[ref.data$density==0.0]<-0.0001
ref.data$density[ref.data$density==1.0]<-0.9999

density.obs <- coefficients(betareg(density ~ group_with_thkey, data = ref.data))[2]



```


```{r}



## information transfer
###### average path length ######

range(ref.data$avg_path_length, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=avg_path_length)) +
  geom_histogram()  # very right skewed

  # shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$avg_path_length)


#m <- lm(avg_path_length ~ withkey, data = ref.data)
m<- glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$avg_path_length)
fitdistrplus::descdist(ref.data$avg_path_length, discrete = F) # beta distribution

ggplot(ref.data, aes(x=group_with_key, y=avg_path_length)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$avg_path_length, proxXkey$withkey)
wilcox.test(avg_path_length~group_with_key, data=ref.data)

avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_thkey, data = ref.data, family=Gamma("inverse")))[2]
m1<- glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse"))
avg_path_length.obs2 <- coefficients(lm(avg_path_length ~ group_with_key, data = ref.data))[2]
m2 <- lm(avg_path_length ~ group_with_key, data = ref.data)

performance::compare_performance(m1,m2, rank = TRUE, verbose = FALSE)
performance::check_model(m1)
performance::check_model(m2)



```

```{r}


##### efficiency #####
range(ref.data$efficiency, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=efficiency)) +
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$efficiency)


m <- lm(efficiency ~ group_with_key, data = ref.data)
#m<- glm(efficiency~ withkey, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
  fitdistrplus::plotdist(ref.data$efficiency)
fitdistrplus::descdist(ref.data$efficiency, discrete = F)


# Calculate effects
ref.data$efficiency[ref.data$efficiency==0.0]<-0.0001
ref.data$efficiency[ref.data$efficiency==1.0]<-0.9999

efficiency.obs <- coefficients(betareg(efficiency ~ group_with_thkey, data = ref.data))[2]

ggplot(ref.data, aes(x=group_with_key, y=efficiency)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$efficiency~proxXkey$withkey)
wilcox.test(efficiency~group_with_key, data=ref.data)


```

```{r}

## combine observed coefficients
network <- "affiliative"
aff.obs<- cbind.data.frame(network, count.obs, diameter.obs,density.obs, avg_path_length.obs, efficiency.obs)


#observed.summary <- rbind(prox.obs, aff.obs, agg.obs)

#observed.summary <- rbind(observed.summary, coefficients.obs)
#}


# 
affXkey <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") 

# count
ggplot(affXkey, aes(x=withkey, y=count)) +
  geom_boxplot() +
  theme_classic()

## connectivity
# diameter
ggplot(affXkey, aes(x=withkey, y=diameter)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(affXkey$diameter, affXkey$withkey)
wilcox.test(diameter~withkey, data=affXkey)

# density
ggplot(affXkey, aes(x=withkey, y=density)) +
  geom_boxplot() +
  theme_classic()

wilcox.test(density~withkey, data=affXkey)
#kruskal.test(affXkey$density, affXkey$withkey)


## information transfer
# average path length
ggplot(affXkey, aes(x=withkey, y=avg_path_length)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(affXkey$avg_path_length, affXkey$withkey)
wilcox.test(avg_path_length~withkey, data=affXkey)

# efficiency
ggplot(affXkey, aes(x=withkey, y=efficiency)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(affXkey$efficiency~affXkey$withkey)
wilcox.test(efficiency~withkey, data=affXkey)
```



### observed agonistic
```{r}

ref.data <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") 

```

```{r number of associations}

range(ref.data$count)

# normally distributed?
ggplot(ref.data, aes(x=count)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed


# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$count)


m <- lm(count ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$count)
fitdistrplus::descdist(ref.data$count, discrete = T) # normal distribution


ggplot(ref.data, aes(x=group_with_key, y=count)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$diameter, proxXkey$withkey)
wilcox.test(count~group_with_key, data=ref.data)
# Calculate effects
count.obs <- coefficients(glm(count ~ group_with_thkey, data = ref.data, family = poisson))[2]
#summary(glm(count ~ group_with_key, data = ref.data, family = poisson))

```


```{r connectivity}

###### connectivity

range(ref.data$diameter)

# normally distributed?
ggplot(ref.data, aes(x=diameter)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$diameter)


m <- lm(diameter ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$diameter)
fitdistrplus::descdist(ref.data$diameter, discrete = F) # beta distribution


ggplot(ref.data, aes(x=group_with_key, y=diameter)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$diameter, proxXkey$withkey)
wilcox.test(diameter~group_with_key, data=ref.data)
# Calculate effects
diameter.obs <- coefficients(lm(diameter ~ group_with_thkey, data = ref.data))[2]


```

```{r density}


####### density  #####

range(ref.data$density)

# normally distributed?
ggplot(ref.data, aes(x=density)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$density)


m <- lm(density ~ group_with_key, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$density)
fitdistrplus::descdist(ref.data$density, discrete = F) # beta distribution

ggplot(ref.data, aes(x=group_with_key, y=density)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$density, proxXkey$withkey)
wilcox.test(density~group_with_key, data=ref.data)

# Calculate effects
ref.data$density[ref.data$density==0.0]<-0.0001
ref.data$density[ref.data$density==1.0]<-0.9999

density.obs <- coefficients(betareg(density ~ group_with_thkey, data = ref.data))[2]



```


```{r}
## information transfer
###### average path length ######

range(ref.data$avg_path_length, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=avg_path_length)) +
  geom_histogram()  # very right skewed

  # shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$avg_path_length)


#m <- lm(avg_path_length ~ withkey, data = ref.data)
m<- glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$avg_path_length)
#fitdistrplus::descdist(ref.data$avg_path_length, discrete = F) # beta distribution

ggplot(ref.data, aes(x=group_with_key, y=avg_path_length)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$avg_path_length, proxXkey$withkey)
wilcox.test(avg_path_length~group_with_key, data=ref.data)


m1<- glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse"))
m2 <- lm(avg_path_length ~ group_with_key, data = ref.data)

performance::compare_performance(m1,m2, rank = TRUE, verbose = FALSE)
performance::check_model(m1)
performance::check_model(m2)


avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_thkey, data = ref.data, family=Gamma("inverse")))[2]
#avg_path_length.obs <- coefficients(lm(avg_path_length ~ group_with_key, data = ref.data))[2]


```


```{r}


##### efficiency #####
range(ref.data$efficiency, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=efficiency)) +
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$efficiency)


m <- lm(efficiency ~ group_with_key, data = ref.data)
#m<- glm(efficiency~ withkey, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
  fitdistrplus::plotdist(ref.data$efficiency)
fitdistrplus::descdist(ref.data$efficiency, discrete = F)


# Calculate effects
ref.data$efficiency[ref.data$efficiency==0.0]<-0.0001
ref.data$efficiency[ref.data$efficiency==1.0]<-0.9999

efficiency.obs <- coefficients(betareg(efficiency ~ group_with_thkey, data = ref.data))[2]

ggplot(ref.data, aes(x=group_with_key, y=efficiency)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(proxXkey$efficiency~proxXkey$withkey)
wilcox.test(efficiency~group_with_key, data=ref.data)

```



```{r}
## combine observed coefficients
network <-"agonistic"
agg.obs<- cbind.data.frame(network, count.obs, diameter.obs,density.obs, avg_path_length.obs, efficiency.obs)

#observed.summary <- rbind(prox.obs, aff.obs, agg.obs)

#observed.summary <- rbind(observed.summary, coefficients.obs)
#}

#plots


aggXkey <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") 

## connectivity
# diameter
ggplot(aggXkey, aes(x=group_with_key, y=diameter)) +
  geom_boxplot() +
  theme_classic()

wilcox.test(diameter~group_with_key, data=aggXkey)
#kruskal.test(aggXkey$diameter, aggXkey$withkey)

# density
ggplot(aggXkey, aes(x=group_with_key, y=density)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(aggXkey$density, aggXkey$withkey)
wilcox.test(density~group_with_key, data=aggXkey)

## information transfer
# average path length
ggplot(aggXkey, aes(x=group_with_key, y=avg_path_length)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(aggXkey$avg_path_length, aggXkey$withkey)
wilcox.test(avg_path_length~group_with_key, data=aggXkey)

# efficiency
ggplot(aggXkey, aes(x=group_with_key, y=efficiency)) +
  geom_boxplot() +
  theme_classic()

#kruskal.test(aggXkey$efficiency~aggXkey$withkey)
wilcox.test(efficiency~group_with_key, data=aggXkey)
```


In the agonistic networks, we found no difference in network metrics between groups with key and without key individuals present. 

### observed multilayer network

```{r}
multiplexmetricsXgroup$group <- as.integer(multiplexmetricsXgroup$group )

multiXkey <-  multiplexmetricsXgroup %>% 
  left_join(groups.key)   
```
``` {r connected}

####### density  #####

range(multiXkey$density)

# normally distributed?
ggplot(multiXkey, aes(x=density)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(multiXkey$density)


m <- lm(density ~ group_with_thkey, data = multiXkey)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(multiXkey$density)
fitdistrplus::descdist(multiXkey$density, discrete = F) # beta distribution

ggplot(multiXkey, aes(x=group_with_thkey, y=density)) +
  geom_boxplot() +
  theme_classic()

wilcox.test(density~group_with_thkey, data=multiXkey)

# Calculate effects
multiXkey$density[multiXkey$density==1.0]<-0.9999

#density.obs <- coefficients(betareg(density ~ group_with_key, data = multiXkey))[2]
density.obs <- coefficients(betareg(density ~ group_with_thkey, data = multiXkey))[2]
#density.obs <- coefficients(betareg(density ~ group_with_qkey, data = multiXkey))[2]
```

```{r information transfer}

###### average path length ######
ref.data <- multiXkey %>% 
   filter(m.apl<0.5)

range(ref.data$m.apl, na.rm=T)

# normally distributed?
ggplot(ref.data, aes(x=m.apl)) +
  geom_histogram()  # very right skewed

  # shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$m.apl)


m <- lm(m.apl ~ group_with_thkey, data = ref.data)
#m<- glm(avg_path_length ~ withkey, data = ref.data, family=Gamma("inverse"))
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(ref.data$m.apl)
#fitdistrplus::descdist(ref.data$avg_path_length, discrete = F) # beta distribution

ggplot(ref.data, aes(x=group_with_thkey, y=m.apl)) +
  geom_boxplot() +
  theme_classic()


m1<- glm(m.apl ~ group_with_thkey, data = ref.data, family=Gamma("inverse"))
m2 <- lm(m.apl ~ group_with_thkey, data = ref.data)

performance::compare_performance(m1,m2, rank = TRUE, verbose = FALSE)
performance::check_model(m1)
performance::check_model(m2)


#avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_key, data = ref.data, family=Gamma("inverse")))[2]
avg_path_length.obs <- coefficients(glm(m.apl ~ group_with_thkey, data = ref.data, family=Gamma("inverse")))[2]
#avg_path_length.obs <- coefficients(glm(avg_path_length ~ group_with_qkey, data = ref.data, family=Gamma("inverse")))[2]

#avg_path_length.obs <- coefficients(lm(m.apl ~ group_with_thkey, data = ref.data))[2]


```


```{r cohesion}


range(multiXkey$cluster)

# normally distributed?
ggplot(multiXkey, aes(x=cluster)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(ref.data$cluster)


m <- lm(cluster ~ group_with_thkey, data = ref.data)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(multiXkey$cluster)
fitdistrplus::descdist(multiXkey$cluster, discrete = F) # beta distribution

ggplot(multiXkey, aes(x=group_with_thkey, y=cluster)) +
  geom_boxplot() +
  theme_classic()

cluster.obs <- coefficients(lm(cluster ~ group_with_thkey, data = multiXkey))[2]


```


```{r cohesion}


range(multiXkey$overlap)

ref.data <- multiXkey %>% 
   filter(overlap<0.2)

# normally distributed?
ggplot(ref.data, aes(x=overlap)) +  # BordaScore, bordarankreverse, bordareverse
  geom_histogram()  # very right skewed

# shapiro test p >0.05, data normally distributed
#ggqqplot(trap, "time_s")
shapiro.test(multiXkey$overlap)


m <- lm(overlap ~ group_with_thkey, data = multiXkey)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = stats::rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 258 (time_s =14.6)

# distribution
fitdistrplus::plotdist(multiXkey$overlap)
fitdistrplus::descdist(multiXkey$overlap, discrete = F) # beta distribution

fitdistrplus::plotdist(ref.data$overlap)
fitdistrplus::descdist(ref.data$overlap, discrete = F) # beta distribution

ggplot(multiXkey, aes(x=group_with_thkey, y=overlap)) +
  geom_boxplot() +
  theme_classic()

overlap.obs <- coefficients(lm(overlap ~ group_with_thkey, data = ref.data))[2]
```

```{r}
## combine observed coefficients
network <-"multilayer"
multilayer.obs<- cbind.data.frame(network, 
                                  density.obs, 
                                  avg_path_length.obs, 
                                  cluster.obs,
                                  overlap.obs)
write.csv(multilayer.obs, "../results/observed multilayer metrics groups_threshold key ids.csv")
```



### combine results
```{r}
observed.summary <- bind_rows(prox.obs, aff.obs, agg.obs)

write.csv(observed.summary, "../results/observed summary metrics groups_threshold key ids.csv")
```




##	Reference model
create reference model where group membership is randomized and calculate network metrics and compare 

We excluded solitary degus.  
We have 7 groups with key individuals present and 5 groups without. 


### summarize ref proximity

```{r create ref model}

groupXkey <- networkmetricsXgroup_prox %>% 
  left_join(groups.key) %>% 
  dplyr::select(group, group_with_thkey) %>% 
  group_by(group) %>% slice(1)


head(groups.key)
glimpse(groupXkey)

N<- n_distinct(groupXkey$group)

replicates <- 1000

#run=2

#make empty dataframe to write loop results into
ref.model.prox <- data.frame(group=numeric(),
                       network=character(),
                       count=numeric(),
                       diameter=numeric(),
                       density=numeric(),
                       avg_path_length=numeric(),
                       efficiency=numeric(),
                       group_with_thkey=character(),
                       random.key=character())


for (run in 1:replicates) {
  r.seed <- run
  RNGkind(sample.kind="default") # code to get same random numbers across operating systems (mac vs windows)
  set.seed(r.seed)
  
  
  # Randomise groups with key individuals
	s <- sample(1:N)
	groupXkey$random.key <- groupXkey$group_with_thkey[s]
	                     
  ref.data <- networkmetricsXgroup_prox %>% 
    select(group, network, count, diameter, density, avg_path_length, efficiency) %>% 
  left_join(groupXkey)
  
  ref.data$runID <- rep(paste0("run",str_pad(r.seed, 3, side="left", pad = "0")), length(networkmetricsXgroup_prox$group)) #generates run IDs that are all 3 digits (001-100), change to 4 digit-padding if running >999
    
  ref.data <- as.data.frame(ref.data)
  ref.model.prox <- rbind.data.frame(ref.model.prox, ref.data)
}

#write.csv(ref.model, "results/refmodel_function_fall24.csv")


```


```{r summarize ref model - proximity}

job::job({
run.s <- as.character(unique(ref.model$runID))
replicates <- 1000 #how many runs of the reference model were done



ref.summary.prox<- data.frame(network=character(),
                              run.code=character(),
                              random.key=character(),
                              count=numeric(),
                              diameter.ref=numeric(),
                              density.ref=numeric(),
                              avg_path_length.ref=numeric(),
                              efficiency.ref=numeric())


start.time <- Sys.time()

for(run in 1:replicates){
  run.code <- run.s[run]
  print(run.code)
  
  run.data <- filter(ref.model.prox, 
                     #network=="proximity",
                     runID==run.code)

network <- "proximity"


###### number of associations
count.ref <- coefficients(glm(count ~ random.key, data = run.data, family = poisson))[2]


###### connectivity
diameter.ref <- coefficients(lm(diameter ~ random.key, data = run.data))[2]



####### density  
run.data$density[run.data$density==0.0]<-0.0001
run.data$density[run.data$density==1.0]<-0.9999

density.ref <- coefficients(betareg(density ~ random.key, data = run.data))[2]


## information transfer
###### average path length 

avg_path_length.ref <- coefficients(glm(avg_path_length ~ random.key, data = run.data, family=Gamma("inverse")))[2]


##### efficiency 
run.data$efficiency[run.data$efficiency==0.0]<-0.0001
run.data$efficiency[run.data$efficiency==1.0]<-0.9999

efficiency.ref <- coefficients(betareg(efficiency ~ random.key, data = run.data))[2]



## combine random coefficients

coefficients.ref <- cbind.data.frame(network, 
                                     run.code, 
                                     count.ref, 
                                     diameter.ref,
                                     density.ref, 
                                     avg_path_length.ref, 
                                     efficiency.ref)

ref.summary.prox <- rbind(ref.summary.prox, coefficients.ref)

}
end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.summaries)
head(ref.summary.prox)
tail(ref.summary.prox)

})

# write data
#write.csv(ref.model.summaries, "ANALYZE-SUMM.refmodel1.csv")


```

### summarize ref affiliative

```{r}

networkmetricsXgroup_aff$group <- as.integer(networkmetricsXgroup_aff$group)
groupXkey <- networkmetricsXgroup_aff %>% 
  left_join(groups.key) %>% 
  dplyr::select(group, group_with_thkey) %>% 
  group_by(group) %>% slice(1)


head(groups.key)
glimpse(groupXkey)

N<- n_distinct(groupXkey$group)

replicates <- 1000

#run=2

#make empty dataframe to write loop results into
ref.model.aff <- data.frame(group=numeric(),
                       network=character(),
                       count=numeric(),
                       diameter=numeric(),
                       density=numeric(),
                       avg_path_length=numeric(),
                       efficiency=numeric(),
                       group_with_thkey=character(),
                       random.key=character())

job::job({

for (run in 1:replicates) {
  r.seed <- run
  RNGkind(sample.kind="default") # code to get same random numbers across operating systems (mac vs windows)
  set.seed(r.seed)
  
  
  # Randomise groups with key individuals
	s <- sample(1:N)
	groupXkey$random.key <- groupXkey$group_with_thkey[s]
	                     
  ref.data <- networkmetricsXgroup_aff %>% 
    select(group, network, count, diameter, density, avg_path_length, efficiency) %>% 
  left_join(groupXkey)
  
  ref.data$runID <- rep(paste0("run",str_pad(r.seed, 3, side="left", pad = "0")), length(networkmetricsXgroup_aff$group)) #generates run IDs that are all 3 digits (001-100), change to 4 digit-padding if running >999
    
  ref.data <- as.data.frame(ref.data)
  ref.model.aff <- rbind.data.frame(ref.model.aff, ref.data)
}
  
})
```



```{r summarize ref model - affiliative}
job::job({

run.s <- as.character(unique(ref.model.aff$runID))
replicates <- 1000 #how many runs of the reference model were done


start.time <- Sys.time()

ref.summary.aff <- data.frame(network=character(),
                              run.code=character(),
                              count.ref=numeric(),
                              diameter.ref=numeric(),
                              density.ref=numeric(),
                              avg_path_length.ref=numeric(),
                              efficiency.ref=numeric())


start.time <- Sys.time()

for(run in 1:replicates){
  run.code <- run.s[run]
  print(run.code)
  run.data <- filter(ref.model.aff, 
                     #network=="affiliative",
                     runID==run.code)

network <- "affiliative"

###### number of interactions
count.ref <- coefficients(glm(count ~ random.key, data = run.data, family = poisson))[2]

###### connectivity
diameter.ref <- coefficients(lm(diameter ~ random.key, data = run.data))[2]



####### density  
run.data$density[run.data$density==0.0]<-0.0001
run.data$density[run.data$density==1.0]<-0.9999

density.ref <- coefficients(betareg(density ~ random.key, data = run.data))[2]


## information transfer
###### average path length 

avg_path_length.ref <- coefficients(glm(avg_path_length ~ random.key, data = run.data, family=Gamma("inverse")))[2]


##### efficiency 
run.data$efficiency[run.data$efficiency==0.0]<-0.0001
run.data$efficiency[run.data$efficiency==1.0]<-0.9999

efficiency.ref <- coefficients(betareg(efficiency ~ random.key, data = run.data))[2]



## combine random coefficients

coefficients.ref <- cbind.data.frame(network, 
                                     run.code, 
                                     count.ref,
                                     diameter.ref,
                                     density.ref, 
                                     avg_path_length.ref, 
                                     efficiency.ref)

ref.summary.aff <- rbind(ref.summary.aff, coefficients.ref)

}
end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.summaries)
head(ref.summary.aff)
tail(ref.summary.aff)

})


# write data
#write.csv(ref.model.summaries, "ANALYZE-SUMM.refmodel1.csv")


```
### summarize ref agonistic

```{r}

networkmetricsXgroup_agg$group <- as.integer(networkmetricsXgroup_agg$group)
groupXkey.agg <- networkmetricsXgroup_agg %>% 
  left_join(groups.key) %>% 
  dplyr::select(group, group_with_thkey) %>% 
  group_by(group) %>% slice(1)


N<- n_distinct(groupXkey.agg$group)

replicates <- 1000

#run=2

#make empty dataframe to write loop results into
ref.model.agg <- data.frame(group=numeric(),
                       network=character(),
                       count=numeric(),
                       diameter=numeric(),
                       density=numeric(),
                       avg_path_length=numeric(),
                       efficiency=numeric(),
                       group_with_thkey=character(),
                       random.key=character())

job::job({

for (run in 1:replicates) {
  r.seed <- run
  RNGkind(sample.kind="default") # code to get same random numbers across operating systems (mac vs windows)
  set.seed(r.seed)
  
  
  # Randomise groups with key individuals
	s <- sample(1:N)
	groupXkey.agg$random.key <- groupXkey.agg$group_with_thkey[s]
	                     
  ref.data <- networkmetricsXgroup_agg %>% 
    select(group, network, count, diameter, density, avg_path_length, efficiency) %>% 
  left_join(groupXkey.agg)
  
  ref.data$runID <- rep(paste0("run",str_pad(r.seed, 3, side="left", pad = "0")), length(networkmetricsXgroup_agg$group)) #generates run IDs that are all 3 digits (001-100), change to 4 digit-padding if running >999
    
  ref.data <- as.data.frame(ref.data)
  ref.model.agg <- rbind.data.frame(ref.model.agg, ref.data)
}
  
})
```


```{r summarize ref model - agonsitic}

job::job({
run.s <- as.character(unique(ref.model$runID))
replicates <- 1000 #how many runs of the reference model were done


start.time <- Sys.time()

ref.summary.ago<- data.frame(network=character(),
                             run.code=character(),
                             count.ref=numeric(),
                             diameter.ref=numeric(),
                             density.ref=numeric(),
                             avg_path_length.ref=numeric(),
                             efficiency.ref=numeric())


start.time <- Sys.time()

for(run in 1:replicates){
  run.code <- run.s[run]
  print(run.code)
  run.data <- filter(ref.model.agg, 
                     #network=="agonistic",
                     runID==run.code)

network <- "agonistic"

###### number of associations
count.ref <- coefficients(glm(count ~ random.key, data = run.data, family = poisson))[2]

###### connectivity
diameter.ref <- coefficients(lm(diameter ~ random.key, data = run.data))[2]


####### density  
run.data$density[run.data$density==0.0]<-0.0001
run.data$density[run.data$density==1.0]<-0.9999

density.ref <- coefficients(betareg(density ~ random.key, data = run.data))[2]


## information transfer
###### average path length 

avg_path_length.ref <- coefficients(glm(avg_path_length ~ random.key, data = run.data, family=Gamma("inverse")))[2]


##### efficiency 
run.data$efficiency[run.data$efficiency==0.0]<-0.0001
run.data$efficiency[run.data$efficiency==1.0]<-0.9999

efficiency.ref <- coefficients(betareg(efficiency ~ random.key, data = run.data))[2]



## combine random coefficients

coefficients.ref <- cbind.data.frame(network, run.code, 
                                     count.ref,
                                     diameter.ref,
                                     density.ref, 
                                     avg_path_length.ref, 
                                     efficiency.ref)

ref.summary.ago <- rbind(ref.summary.ago, coefficients.ref)

}
end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.summaries)
head(ref.summary.ago)
tail(ref.summary.ago)

})


# write data
#write.csv(ref.model.summaries, "ANALYZE-SUMM.refmodel1.csv")


```


###summarize ref multilayer

```{r}

networkmetricsXgroup$group <- as.integer(networkmetricsXgroup_agg$group)
groupXkey.multi <- multiplexmetricsXgroup %>% 
  left_join(groups.key) %>% 
  dplyr::select(group, group_with_thkey) %>% 
  group_by(group) %>% slice(1)


N<- n_distinct(groupXkey.multi$group)

replicates <- 1000

#run=2

#make empty dataframe to write loop results into
ref.model.multi <- data.frame(group=numeric(),
                       network=character(),
                       diameter=numeric(),
                       density=numeric(),
                       avg_path_length=numeric(),
                       cluster=numeric(),
                       overlap=numeric(),
                       group_with_thkey=character(),
                       random.key=character())

job::job({

for (run in 1:replicates) {
  r.seed <- run
  RNGkind(sample.kind="default") # code to get same random numbers across operating systems (mac vs windows)
  set.seed(r.seed)
  
  
  # Randomise groups with key individuals
	s <- sample(1:N)
	groupXkey.multi$random.key <- groupXkey.multi$group_with_thkey[s]
	                     
  ref.data <- multiplexmetricsXgroup %>% 
    select(group, network, diameter, density, m.apl, cluster, overlap) %>% 
  left_join(groupXkey.multi)
  
  ref.data$runID <- rep(paste0("run",str_pad(r.seed, 3, side="left", pad = "0")), length(multiplexmetricsXgroup$group)) #generates run IDs that are all 3 digits (001-100), change to 4 digit-padding if running >999
    
  ref.data <- as.data.frame(ref.data)
  ref.model.multi <- rbind.data.frame(ref.model.multi, ref.data)
}
  
})
```


```{r summarize ref model - multilayer}

job::job({
run.s <- as.character(unique(ref.model.multi$runID))
replicates <- 1000 #how many runs of the reference model were done


start.time <- Sys.time()

ref.summary.multi<- data.frame(network=character(),
                             run.code=character(),
                             #diameter.ref=numeric(),
                             density.ref=numeric(),
                             avg_path_length.ref=numeric(),
                             cluster.ref=numeric(),
                             overlap.ref=numeric())


for(run in 1:replicates){
  run.code <- run.s[run]
  print(run.code)
  run.data <- filter(ref.model.multi, 
                     #network=="agonistic",
                     runID==run.code)

network <- "multilayer"


####### density  
run.data$density[run.data$density==0.0]<-0.0001
run.data$density[run.data$density==1.0]<-0.9999

density.ref <- coefficients(betareg(density ~ random.key, data = run.data))[2]


## information transfer
###### average path length 

avg_path_length.ref <- coefficients(glm(m.apl ~ random.key, data = run.data, family=Gamma("inverse")))[2]


# cohesion
cluster.ref <- coefficients(lm(cluster ~ random.key, data = run.data))[2]
overlap.ref <- coefficients(lm(overlap ~ random.key, data = run.data))[2]

## combine random coefficients

coefficients.ref <- cbind.data.frame(network, run.code, 
                                     #count.ref,
                                     #diameter.ref,
                                     density.ref, 
                                     avg_path_length.ref, 
                                     cluster.ref,
                                     overlap.ref)

ref.summary.multi <- rbind(ref.summary.multi, coefficients.ref)

}
end.time <- Sys.time()

#Time to run:
end.time - start.time

#t(ref.model.summaries)
head(ref.summary.multi)
tail(ref.summary.multi)

})


# write data
#write.csv(ref.model.summaries, "ANALYZE-SUMM.refmodel1.csv")


```




## Visualise results

### proximity
 Are observed values significantly different from random values in reference models?

find proportion of random values less than the observed values, you can use this as a kind of p value. p should be >0.975 for two-tailed test, >0.95 for one-tailed test for something significantly different from the random values in the reference model
  

```{r proximity results}
glimpse(observed.summary)
glimpse(prox.obs)
glimpse(ref.summary.prox)


sum(prox.obs$count.obs[prox.obs$network=="proximity"]<ref.summary.prox$count.ref)/length(ref.summary.prox$count.ref)

plot_prox_count_p <- ggplot(ref.summary.prox, aes(x=count.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Count coefficient value ", y="Frequency", title = "P = 0.64") +
  #geom_vline(xintercept=observed.summary$diameter.obs[observed.summary$network=="proximity"], color="red") +
  geom_vline(xintercept=prox.obs$count.obs[prox.obs$network=="proximity"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$count.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$count.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-0.135, y=81, label="p = 0.53")
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_prox_count <- networkmetricsXgroupXkey %>% 
  filter(network=="proximity") %>%
  ggplot( aes(x=group_with_thkey, y=count, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Count", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 




1-sum(prox.obs$diameter.obs[prox.obs$network=="proximity"]<ref.summary.prox$diameter.ref)/length(ref.summary.prox$diameter.ref)

plot_prox_diameter <- ggplot(ref.summary.prox, aes(x=diameter.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Diameter coefficient value ", y="Frequency", title = "P = 0.01") +
  #geom_vline(xintercept=observed.summary$diameter.obs[observed.summary$network=="proximity"], color="red") +
  geom_vline(xintercept=prox.obs$diameter.obs[prox.obs$network=="proximity"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$diameter.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$diameter.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-0.135, y=81, label="p = 0.53")
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_prox_dia <- networkmetricsXgroupXkey %>% 
  filter(network=="proximity") %>%
  ggplot( aes(x=group_with_thkey, y=diameter, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Diameter", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

 sum(prox.obs$density.obs[prox.obs$network=="proximity"]<ref.summary.prox$density.ref)/length(ref.summary.prox$density.ref)
plot_prox_density <-  ggplot(ref.summary.prox, aes(x=density.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Density coefficient value", y="Frequency", title = "P = 0.17") +
  geom_vline(xintercept=prox.obs$density.obs[prox.obs$network=="proximity"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$density.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$density.ref, 0.975), 2), color="darkblue", lty=2) +
  #  geom_text(x=-2.7, y=225, label="p = 0.46") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
 
plot_prox_den <- networkmetricsXgroupXkey %>% 
  filter(network=="proximity") %>%
  ggplot( aes(x=group_with_thkey, y=density, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Density", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 


sum(prox.obs$avg_path_length.obs[prox.obs$network=="proximity"]<ref.summary.prox$avg_path_length.ref)/length(ref.summary.prox$avg_path_length.ref)
plot_prox_apl_p <- ggplot(ref.summary.prox, aes(x=avg_path_length.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Average path length coefficient value", y="Frequency", title="P = 0.23") +
  geom_vline(xintercept=prox.obs$avg_path_length.obs[prox.obs$network=="proximity"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$avg_path_length.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$avg_path_length.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-4.5, y=80, label="p = 0.42") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_prox_apl <- networkmetricsXgroupXkey %>% 
  filter(network=="proximity") %>%
  ggplot( aes(x=group_with_thkey, y=avg_path_length, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Average path length", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 
  
   sum(prox.obs$efficiency.obs[prox.obs$network=="proximity"]<ref.summary.prox$efficiency.ref)/length(ref.summary.prox$efficiency.ref)
plot_prox_eff_p <-   ggplot(ref.summary.prox, aes(x=efficiency.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Efficiency coefficient value", y="Frequency", title="P = 0.25") +
  geom_vline(xintercept=prox.obs$efficiency.obs[prox.obs$network=="proximity"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$efficiency.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.prox$efficiency.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-1.5, y=195, label="p = 0.26") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_prox_eff <- networkmetricsXgroupXkey %>% 
  filter(network=="proximity") %>%
  ggplot( aes(x=group_with_thkey, y=efficiency, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Efficiency", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 
   
```

```{r}
ggpubr::ggarrange(plot_prox_count, plot_prox_count_p,
                  plot_prox_dia, plot_prox_diameter,
                 
                  plot_prox_den,  plot_prox_density, 
                  
                  plot_prox_apl,plot_prox_apl_p,
                  
                  plot_prox_eff,plot_prox_eff_p,
                  nrow = 5, ncol=2)

# Add the inset plot to the main plot
plot_prox_cou1 <- plot_prox_count +
  annotation_custom(
    grob = ggplotGrob(plot_prox_count_p ),
    xmin = 1.2, xmax = 1.8,
    ymin = 300, ymax = 450
  )

plot_prox_dia1 <- plot_prox_dia +
  annotation_custom(
    grob = ggplotGrob(plot_prox_diameter ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.8,
    ymin = 1.0, ymax = 1.5
  )

plot_prox_den1 <- plot_prox_den +
  annotation_custom(
    grob = ggplotGrob(plot_prox_density ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.85,
    ymin = 0.7, ymax = 0.85
  )

plot_prox_apl1 <- plot_prox_apl +
  annotation_custom(
    grob = ggplotGrob(plot_prox_apl_p ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.9,
    ymin = 0.6, ymax = 0.9
  )

plot_prox_eff1 <- plot_prox_eff +
  annotation_custom(
    grob = ggplotGrob(plot_prox_eff_p ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.9,
    ymin = 0.6, ymax = 1
  )



ggpubr::ggarrange(plot_prox_cou1 + rremove('xlab') , 
                  plot_prox_dia1 + rremove('xlab') , 
                  plot_prox_den1 + rremove('xlab'),
                  plot_prox_apl1, 
                  plot_prox_eff1
                  )
ggsave("figures/groupmetrics_prox_fall24.pdf",width=180, height = 120, units = "mm", dpi = 300)
```



### affiliative

```{r affiliative results}
glimpse(ref.summary.aff)

sum(observed.summary$count.obs[observed.summary$network=="affiliative"]<ref.summary.aff$count.ref)/length(ref.summary.aff$count.ref)
plot_aff_count_p <-  ggplot(ref.summary.aff, aes(x=count.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Count coefficient value", y="Frequency", title="P = 0.69") +
  geom_vline(xintercept=observed.summary$count.obs[observed.summary$network=="affiliative"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$count.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$count.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-0.135, y=85, label="p = 0.64") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_aff_count <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") %>%
  ggplot( aes(x=group_with_thkey, y=count, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Count", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

sum(observed.summary$diameter.obs[observed.summary$network=="affiliative"]<ref.summary.aff$diameter.ref)/length(ref.summary.aff$diameter.ref)
plot_aff_diameter <-  ggplot(ref.summary.aff, aes(x=diameter.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Diameter coefficient value", y="Frequency", title="P = 0.51") +
  geom_vline(xintercept=observed.summary$diameter.obs[observed.summary$network=="affiliative"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$diameter.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$diameter.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-0.135, y=85, label="p = 0.64") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_aff_dia <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") %>%
  ggplot( aes(x=group_with_thkey, y=diameter, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Diameter", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 


 sum(observed.summary$density.obs[observed.summary$network=="affiliative"]<ref.summary.aff$density.ref)/length(ref.summary.aff$density.ref)
plot_aff_density <- ggplot(ref.summary.aff, aes(x=density.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Density coefficient value", y="Frequency", title="P = 0.06") +
  geom_vline(xintercept=observed.summary$density.obs[observed.summary$network=="affiliative"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$density.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$density.ref, 0.975), 2), color="darkblue", lty=2) +
    #geom_text(x=-2.7, y=225, label="") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
 
plot_aff_den <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") %>%
  ggplot( aes(x=group_with_thkey, y=density, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Density", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

  sum(observed.summary$avg_path_length.obs[observed.summary$network=="affiliative"]<ref.summary.aff$avg_path_length.ref)/length(ref.summary.aff$avg_path_length.ref)
 plot_aff_apl_p <- ggplot(ref.summary.aff, aes(x=avg_path_length.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Average path length coefficient value", y="Frequency", title="P = 0.48") +
  geom_vline(xintercept=observed.summary$avg_path_length.obs[observed.summary$network=="affiliative"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$avg_path_length.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$avg_path_length.ref, 0.975), 2), color="darkblue", lty=2) +
     #geom_text(x=-4.5, y=70, label="") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
 
 plot_aff_apl <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") %>%
  ggplot( aes(x=group_with_thkey, y=avg_path_length, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Average path length", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 
  
  
   sum(observed.summary$efficiency.obs[observed.summary$network=="affiliative"]<ref.summary.aff$efficiency.ref)/length(ref.summary.aff$efficiency.ref)
plot_aff_eff_p <- ggplot(ref.summary.aff, aes(x=efficiency.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Efficiency coefficient value", y="Frequency", title="P = 0.01") +
  geom_vline(xintercept=observed.summary$efficiency.obs[observed.summary$network=="affiliative"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$efficiency.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.aff$efficiency.ref, 0.975), 2), color="darkblue", lty=2)  +
      #  geom_text(x=-0.4, y=81, label="p = 0.53") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

  plot_aff_eff <- networkmetricsXgroupXkey %>% 
  filter(network=="affiliative") %>%
  ggplot( aes(x=group_with_thkey, y=efficiency, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Efficiency", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 
 
```

```{r}
ggpubr::ggarrange(plot_aff_count ,plot_aff_count_p , 
                  plot_aff_dia, plot_aff_diameter,
                 
                  plot_aff_den,  plot_aff_density, 
                  
                  plot_aff_apl,plot_aff_apl_p,
                  
                  plot_aff_eff,plot_aff_eff_p,
                  nrow = 5, ncol=2)


# Add the inset plot to the main plot
plot_aff_cou1 <- plot_aff_count +
  annotation_custom(
    grob = ggplotGrob(plot_aff_count_p ),
    xmin = 1.2, xmax = 1.8,
    ymin = 100, ymax = 150
  )

plot_aff_dia1 <- plot_aff_dia +
  annotation_custom(
    grob = ggplotGrob(plot_aff_diameter ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.85,
    ymin = 0.25, ymax = 0.35
  )

plot_aff_den1 <- plot_aff_den +
  annotation_custom(
    grob = ggplotGrob(plot_aff_density ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.85,
    ymin = 0.25, ymax = 0.5
  )

plot_aff_apl1 <- plot_aff_apl +
  annotation_custom(
    grob = ggplotGrob(plot_aff_apl_p ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.85,
    ymin = 0.15, ymax = 0.25
  )

plot_aff_eff1 <- plot_aff_eff +
  annotation_custom(
    grob = ggplotGrob(plot_aff_eff_p ),
   xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.85,
    ymin = 0.01, ymax = 0.06
  )



ggpubr::ggarrange(plot_aff_cou1 + rremove('xlab') , 
                  plot_aff_dia1 + rremove('xlab') , 
                  plot_aff_den1 + rremove('xlab'),
                  plot_aff_apl1, 
                  plot_aff_eff1
                  )
ggsave("figures/groupmetrics_affiliative_fall24.pdf",width=180, height = 120, units = "mm", dpi = 300)

```

### agonistic
```{r agonistic results}
sum(observed.summary$count.obs[observed.summary$network=="agonistic"]<ref.summary.ago$count.ref)/length(ref.summary.ago$count.ref)
plot_agg_count_p <- ggplot(ref.summary.ago, aes(x=count.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Count coefficient value", y="Frequency", title="P = 0.22") +
  geom_vline(xintercept=observed.summary$count.obs[observed.summary$network=="agonistic"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$count.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$count.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-0.03, y=70, label="p = 0.33", size=6) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
        

plot_agg_count <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=count, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Count", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

sum(observed.summary$diameter.obs[observed.summary$network=="agonistic"]<ref.summary.ago$diameter.ref)/length(ref.summary.ago$diameter.ref)
plot_agg_diameter <- ggplot(ref.summary.ago, aes(x=diameter.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Diameter coefficient value", y="Frequency", title="P = 0.15") +
  geom_vline(xintercept=observed.summary$diameter.obs[observed.summary$network=="agonistic"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$diameter.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$diameter.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-0.03, y=70, label="p = 0.33", size=6) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
        

plot_agg_dia <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=diameter, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Diameter", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 


 
sum(observed.summary$density.obs[observed.summary$network=="agonistic"]<ref.summary.ago$density.ref)/length(ref.summary.ago$density.ref)
plot_agg_density <-  ggplot(ref.summary.ago, aes(x=density.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Density coefficient value", y="Frequency", title="P = 0.22") +
  geom_vline(xintercept=observed.summary$density.obs[observed.summary$network=="agonistic"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$density.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$density.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-2, y=125, label="p = 0.76") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
 
plot_agg_den <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=density, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Density", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

  sum(observed.summary$avg_path_length.obs[observed.summary$network=="agonistic"]<ref.summary.ago$avg_path_length.ref)/length(ref.summary.ago$avg_path_length.ref)
plot_agg_apl_p <- ggplot(ref.summary.ago, aes(x=avg_path_length.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Average path length coefficient value", y="Frequency", title="P = 0.81") +
  geom_vline(xintercept=observed.summary$avg_path_length.obs[observed.summary$network=="agonistic"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$avg_path_length.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$avg_path_length.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-11, y=80, label="p = 0.49") +
    theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_agg_apl <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=avg_path_length, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Average path length", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 
  
  
   sum(observed.summary$efficiency.obs[observed.summary$network=="agonistic"]<ref.summary.ago$efficiency.ref)/length(ref.summary.ago$efficiency.ref)
 plot_agg_eff_p <-  ggplot(ref.summary.ago, aes(x=efficiency.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Efficiency coefficient value", y="Frequency", title="P = 0.37") +
  geom_vline(xintercept=observed.summary$efficiency.obs[observed.summary$network=="agonistic"], color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$efficiency.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.ago$efficiency.ref, 0.975), 2), color="darkblue", lty=2) +
      #geom_text(x=-0.7, y=220, label="p = 0.22") +
     theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

   plot_agg_eff <- networkmetricsXgroupXkey %>% 
  filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=efficiency, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Efficiency", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

```




```{r}
ggpubr::ggarrange(plot_agg_count, plot_agg_count_p,
                  plot_agg_dia, plot_agg_diameter,
                 
                  plot_agg_den,  plot_agg_density, 
                  
                  plot_agg_apl,plot_agg_apl_p,
                  
                  plot_agg_eff,plot_agg_eff_p,
                  nrow = 5, ncol=2)



# Add the inset plot to the main plot
plot_agg_cou1 <- plot_agg_count +
  annotation_custom(
    grob = ggplotGrob(plot_agg_count_p ),
    xmin = 1.2, xmax = 1.8,
    ymin = 5, ymax = 7.5
  )

plot_agg_dia1 <- plot_agg_dia +
  annotation_custom(
    grob = ggplotGrob(plot_agg_diameter ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.99,
    ymin = 0.15, ymax = 0.22
  )

plot_agg_den1 <- plot_agg_den +
  annotation_custom(
    grob = ggplotGrob(plot_agg_density ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.9,
    ymin = 0.29, ymax = 0.49
  )

plot_agg_apl1 <- plot_agg_apl +
  annotation_custom(
    grob = ggplotGrob(plot_agg_apl_p ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.99,
    ymin = 0.07, ymax = 0.1
  )

plot_agg_eff1 <- plot_agg_eff +
  annotation_custom(
    grob = ggplotGrob(plot_agg_eff_p ),
    xmin = 1.2, xmax = 1.8,
    #xmin = 0.4, xmax = 0.9,
    ymin = 0.15, ymax = 0.23
  )



ggpubr::ggarrange(plot_agg_cou1 + rremove('xlab') , 
                  plot_agg_dia1 + rremove('xlab') , 
                  plot_agg_den1 + rremove('xlab'),
                  plot_agg_apl1, 
                  plot_agg_eff1
                  )
ggsave("figures/groupmetrics_agonistic_fall24.pdf",width=180, height = 120, units = "mm", dpi = 300)


```


###multilayer

```{r}

# density
 
sum(multilayer.obs$density.obs<ref.summary.multi$density.ref)/length(ref.summary.multi$density.ref)

plot_multi_density <-  ggplot(ref.summary.multi, aes(x=density.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Density coefficient value", y="Frequency", title="P = 0.02") +
  geom_vline(xintercept=multilayer.obs$density.obs, color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$density.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$density.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-2, y=125, label="p = 0.76") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))
 
plot_multi_den <- multiXkey %>% 
  #filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=density, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Density", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 


# average path length
  sum(multilayer.obs$avg_path_length.obs<ref.summary.multi$avg_path_length.ref)/length(ref.summary.multi$avg_path_length.ref)

  plot_multi_apl_p <- ggplot(ref.summary.multi, aes(x=avg_path_length.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Average path length coefficient value", y="Frequency", title="P = 0.35") +
  geom_vline(xintercept=multilayer.obs$avg_path_length.obs, color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$avg_path_length.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$avg_path_length.ref, 0.975), 2), color="darkblue", lty=2) +
  #geom_text(x=-11, y=80, label="p = 0.49") +
    theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_multi_apl <- multiXkey %>% 
  ggplot( aes(x=group_with_thkey, y=m.apl, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Average path length", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

# cluster
sum(multilayer.obs$cluster.obs<ref.summary.multi$cluster.ref)/length(ref.summary.multi$cluster.ref)
 plot_multi_clus_p <-  ggplot(ref.summary.multi, aes(x=cluster.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Cluster coefficient value", y="Frequency", title="P = 0.97") +
  geom_vline(xintercept=multilayer.obs$cluster.obs, color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$cluster.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$cluster.ref, 0.975), 2), color="darkblue", lty=2) +
      #geom_text(x=-0.7, y=220, label="p = 0.22") +
     theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_multi_clus <- multiXkey %>% 
  #filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=cluster, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Cluster", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 
   
   
   # overlap
 sum(multilayer.obs$overlap.obs<ref.summary.multi$overlap.ref)/length(ref.summary.multi$overlap.ref)
 plot_multi_ove_p <-  ggplot(ref.summary.multi, aes(x=overlap.ref)) +
  geom_histogram(color="grey", fill="grey") +
  theme_classic() +
  labs(x= "Overlap coefficient value", y="Frequency", title="P = 0.53") +
  geom_vline(xintercept=multilayer.obs$overlap.obs, color="red") +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$overlap.ref, 0.025), 2), color="darkblue", lty=2) +
  geom_vline(xintercept = rep(quantile(ref.summary.multi$overlap.ref, 0.975), 2), color="darkblue", lty=2) +
      #geom_text(x=-0.7, y=220, label="p = 0.22") +
     theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        #axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        title = element_text(size=6))

plot_multi_ove <- multiXkey %>% 
  #filter(network=="agonistic") %>%
  ggplot( aes(x=group_with_thkey, y=overlap, color=group_with_thkey)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Overlap", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none")   
```
```{r}
ggpubr::ggarrange(plot_multi_den,  plot_multi_density, 
                  
                  plot_multi_apl,plot_multi_apl_p,
                  
                  plot_multi_clus,plot_multi_clus_p,
                  plot_multi_ove,plot_multi_ove_p,
                  nrow = 5, ncol=2)
```



### combined all

```{r only monolayers}
plot_networkmetricsXgroup24 <- ggpubr::ggarrange(plot_prox_count + rremove('xlab'), 
                  plot_aff_count + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_count + rremove('xlab') + rremove('ylab'), 
  
                  plot_prox_dia + rremove('xlab'), 
                  plot_aff_dia + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_dia + rremove('xlab') + rremove('ylab'), 
                  
                  plot_prox_den + rremove('xlab'), 
                  plot_aff_den + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_den + rremove('xlab')+ rremove('ylab'),
                  
                  plot_prox_apl + rremove('xlab'), 
                  plot_aff_apl + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_apl + rremove('xlab')+ rremove('ylab'),
                  
                  plot_prox_eff, 
                  plot_aff_eff+ rremove('ylab'), 
                  plot_agg_eff+ rremove('ylab'),
                  
                  nrow=5, ncol=3)

plot_networkmetricsXgroup24
ggsave("../figures/thkeyXgroupXmetrics_fall24.pdf",width=180, height = 210, units = "mm", dpi = 300)                  
```


```{r all}
plot_networkmetricsXgroup24_all <- ggpubr::ggarrange(plot_prox_count + rremove('xlab'), 
                  plot_aff_count + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_count + rremove('xlab') + rremove('ylab'), 
                  plot_agg_count + rremove('xlab') + rremove('ylab'), # I will have to remove this later
  
                  plot_prox_dia + rremove('xlab'), 
                  plot_aff_dia + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_dia + rremove('xlab') + rremove('ylab'), 
                  plot_multi_clus + rremove('xlab')+ rremove('ylab'),  # cluster instead of diameter
                  
                  plot_prox_den + rremove('xlab'), 
                  plot_aff_den + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_den + rremove('xlab')+ rremove('ylab'),
                  plot_multi_den + rremove('xlab')+ rremove('ylab'),  # density is denisty
                  
                  plot_prox_apl + rremove('xlab'), 
                  plot_aff_apl + rremove('xlab')+ rremove('ylab'), 
                  plot_agg_apl + rremove('xlab')+ rremove('ylab'),
                  plot_multi_apl + rremove('xlab')+ rremove('ylab'),  # apl = apl
                  
                  plot_prox_eff, 
                  plot_aff_eff+ rremove('ylab'), 
                  plot_agg_eff+ rremove('ylab'),
                  plot_multi_ove + rremove('ylab'),  # overlap instead of efficiency
                  
                  nrow=5, ncol=4)

plot_networkmetricsXgroup24_all
ggsave("../figures/thkeyXgroupXmetrics_fall24_all.pdf",width=180, height = 210, units = "mm", dpi = 300)                  
```




# Does group characteristics predict presence of key degus?

```{r}
load("../results/key_roleXgroup_fall2024.Rdata")
```

```{r}
keysex24 <- degulist24 %>% 
  filter(rank<16) %>% 
  select(sex, group) %>% 
  rename(sex.keyID=sex)
```


```{r}
groupchars24 <- degulist24 %>% 
  group_by(group, sex) %>% 
  tally() %>% 
  pivot_wider(#ids_col = group,
              names_from = sex,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sexratio = M/(F+M),
         groupsize=sum(F+M), 
         groups.with.key=if_else(group %in% groups.with.thkey, "yes", "no"),
         key_num=if_else(groups.with.key=="yes", 1, 0)) %>% 
  filter(group!="NA") %>% 
  left_join(keysex24, by="group") %>% 
  mutate(M_key=if_else(sex.keyID=="M", "yes", "no"), 
         F_key=if_else(sex.keyID!="M", "yes", "no"))

groupchars24$M_key[is.na(groupchars24$M_key)] <- "no"
groupchars24$F_key[is.na(groupchars24$F_key)] <- "no"



```


## social organization

If the LRT is significant, it means that adding  a predictor improves your model's fit in a statistically meaningful way



```{r}


sum_groupsize <- groupchars %>% 
  ungroup() %>% 
  summarize(mean=mean(groupsize),
            sd=sd(groupsize),
            n=length(group),
            se=sd/sqrt(n),
            min=min(groupsize),
            max=max(groupsize))

sum_sexratio <- groupchars %>% 
  ungroup() %>% 
  summarize(mean=mean(sexratio),
            sd=sd(sexratio),
            n=length(group),
            se=sd/sqrt(n),
            min=min(sexratio),
            max=max(sexratio))

glimpse(groupchars)


m.all <- glmer( key_num ~ groupsize + sexratio + (1|group),
                  data = groupchars,
                  family = binomial
                  )
summary(m.all)

m.sex <- glmer( key_num ~ sexratio + (1|group),
                  data = groupchars,
                  family = binomial
                  )
summary(m.sex)

m.group <- glmer( key_num ~ groupsize + (1|group),
                  data = groupchars,
                  family = binomial
                  )
summary(m.group)

performance::model_performance(m.all, m.sex, m.group)

lmtest::lrtest(m.all, m.sex)
lmtest::lrtest(m.all, m.group)



plot_groupsize <- ggplot(groupchars,  aes(x=groups.with.key, y=groupsize, color=groups.with.key)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Group size", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

plot_sexratio <-   ggplot(groupchars,  aes(x=groups.with.key, y=sexratio, color=groups.with.key)) +
   geom_violin(width=0.5, color="black")+
  geom_boxplot(width=0.2, color="grey", alpha=0.6) +
  geom_jitter(  width=0.05) +
  theme_classic() +
  scale_color_manual( values=alpha(c("#785EF0","#DC267F"),0.8)) +
  labs(y="Sex ratio", x="Social group with key") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "none") 

ggpubr::ggarrange(plot_groupsize, plot_sexratio)
```

# H2.2 males


2.2. Males are more likely to be key in groups with multiple males than in groups with 0 or 1 male

```{r males}
head(groupchars24)


# Create group factor: "0 or 1 male" and "multiple males"
df24 <- groupchars24 %>%
  mutate(male_group = ifelse(M <= 1, "0 or 1 male", "multiple males"), 
         male_key = ifelse(is_key=="yes", 1, 0))


# Plot the proportion of key males by group type
summary_df24 <- df24 %>%
  group_by(is_key, male_group) %>%
  tally()

ggplot(summary_df24, aes(x=male_group, y=n, fill=is_key)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_classic() +
  #scale_y_reverse() +
  scale_fill_manual(values=alpha(c("#785EF0","#DC267F"),0.8)) + 
  labs( x="Number of males in the group", y="count") +
  theme(#axis.text = element_text(size=18),
        #title = element_text(size=20),
        legend.position = "right") 

```



# Visualize networks
I dont think attributes are correct

Shape: key
circle = key
diamond = nonkey


color fill: sex
red = female
blue = male

Size: rank
bigger nodes = lower ranks (key individuals)

```{r}
library(patchwork) # For plot aggregation
# To aggregate and display all plots together:

clean_plot_list <- Filter(Negate(is.null), plot_prox_list)
proximity_plot <- wrap_plots(clean_plot_list)
#proximity_plot <- cowplot::plot_grid(plotlist = clean_plot_list)
print(proximity_plot)

plot_prox_list[[10]]

```

```{r}
clean_plot_list <- Filter(Negate(is.null), plot_aff_list)
affiliative_plot <- wrap_plots(clean_plot_list)
#affiliative_plot <- cowplot::plot_grid(plotlist = clean_plot_list)
affiliative_plot

plot_aff_list[1]
```


```{r}
clean_plot_list <- Filter(Negate(is.null), plot_agg_list)
agonistic_plot <- wrap_plots(clean_plot_list)
#agonistic_plot <- cowplot::plot_grid(plotlist = clean_plot_list)
agonistic_plot

plot_agg_list[1]
```

# save work

```{r}
save.image("../results/key_roleXgroup_fall2024.Rdata")
```
